---
title: "Programming in Base R"
format: html
---

#Task 1: Basic Vector Practice

1.  Set the vectors for *pre_measures* and *post_measures*:

```{r}
pre_measures <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)
post_measures <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
```

2.  Update the names for each element to *Subject_x*:

```{r}
names(pre_measures) <- paste("Subject", 1:20, sep = "_")
names(post_measures) <- paste("Subject", 1:20, sep = "_")
```

3.  Create another vector representing the difference in the *pre* and *post* measurements for each subject:

```{r}
delta_measures <- post_measures - pre_measures
head(delta_measures)
```

4.  Calculate average decrease in blood pressure over all subjects:

```{r}
mean(delta_measures)
```

5.  Identify the subjects that experienced a positive change (decreased blood pressure measurement):

```{r}
positive_change_indices <- which(delta_measures < 0)
head(positive_change_indices)
```

6.  Now, subset the delta vector for only those with the positive change:

```{r}
positive_change_subjects <- delta_measures[positive_change_indices]
head(positive_change_subjects)
```

7.  Lastly, calculate average decrease in blood pressure *of those who experienced a decrease*:

```{r}
mean(positive_change_subjects)
```

# Task 2: Basic Data Frame Practice

1.  Create a data frame with a column for each of:

-   patient
-   pre_bp
-   post_bp
-   diff_bp

```{r}
bp_df = data.frame(
  "patient" = paste("Subject", 1:20, sep = "_"),
  "pre_bp" = pre_measures,
  "post_bp" = post_measures,
  "diff_bp" = delta_measures,
  row.names = NULL
  )
head(bp_df)
```

2. Access the rows corresponding to negative values of the column *diff_bp*:

```{r}
bp_df[bp_df$diff_bp<0,]
```

3. Add a new column, *TRUE* if post_bp < 120:

```{r}
bp_df$post_bp_under_120 = bp_df$post_bp<120
```

4. Print the finished data frame nicely:

```{r}
knitr::kable(bp_df)
```

# Task 3: List practice

1. Create a new data frame to represent this additional set of subjects that received placebos:

```{r}
placebo_df <- data.frame(
  "patient" = paste("Subject",1:20,sep="_"),
  "pre_bp" = c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130),
  "post_bp" = c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
)

#adding diff and under_120 cols
placebo_df$diff_bp <- placebo_df$post_bp - placebo_df$pre_bp
placebo_df$post_bp_under_120 <- placebo_df$diff_bp<120

head(placebo_df)
```

2. Now make a list whose first element is the original *treatment* group and whose second element is the *placebo* group:

```{r}
patient_list <- list(
  "treatment" = bp_df,
  "placebo" = placebo_df
  )
```

3. Show three ways to access the first element of this list:

```{r}
patient_list[1]
```

```{r}
patient_list[[1]]
```

```{r}
patient_list$treatment
```

4. Access the *pre_bp* column of the *placebo* element with one line of code:

```{r}
patient_list[[2]]$pre_bp
```

